---
title: "Eradication Effectiveness"
author: "Hazel Daniels"
date: "April 24, 2019"
output: html_document
---

# Reading in file 
```{r}
library(readxl)
AllSites <- read_xlsx("C:/Users/dixie/Google Drive/_Ph.D. Forestry/Research/2018.03_EradicationEffectiveness/R Files/BasicData.xlsx")
head(AllSites)
AllSites$VegBin <- as.numeric(as.character(AllSites$VegBin))
AllSites$SoilBin <- as.numeric(as.character(AllSites$SoilBin))
AllSites$TrtBin <- as.numeric(as.character(AllSites$TrtBin))
head(AllSites)
```

# Vegetation
## Fitting a model for veg
```{r}
veg_null = glm(VegBin~1, family = binomial, data=AllSites)
veg_full = glm(VegBin~Lineage * TrtBin * SoilBin, family=binomial, data=AllSites)
step(veg_null, scope = list(upper=veg_full), direction="both", test="Chisq", data=AllSItess)
#VegBin~SoilBin + TrtBin + SoilBin:TrtBin

library(rJava)
library(glmulti)
model_full2 <- glmulti(VegBin ~ Lineage * TrtBin * SoilBin, data=AllSites, crit=bic, method = "g")
#VegBin~1+TrtBin+SoilBin:TrtBin
model_full3 <- glmulti(VegBin ~ Lineage * TrtBin * SoilBin, data=AllSites, crit=aic, method = "g")
#VegBin~1+TrtBin+SoilBin:TrtBin
```

## Fitting a model for NA1 veg only
```{r}
#Subsetting data
NA1Sites <- subset(AllSites, AllSites$Lineage == "NA1")

NAV_null = glm(VegBin~1, family = binomial, data=NA1Sites)
NAV_full = glm(VegBin~TrtBin * SoilBin, family=binomial, data=NA1Sites)
step(NAV_null, scope = list(upper=NAV_full), direction="both", test="Chisq", data=NA1Sites)
#VegBin~SoilBin + TrtBin

NAV_full2 <- glmulti(VegBin ~ TrtBin * SoilBin, data=NA1Sites, crit=bic, method = "g")
#VegBin~1+TrtBin+SoilBin:TrtBin
NAV_full3 <- glmulti(VegBin ~ TrtBin * SoilBin, data=NA1Sites, crit=aic, method = "g")
#VegBin~1+TrtBin+SoilBin:TrtBin
```

## Fitting a model for EU1 veg only
```{r}
#Subsetting data
EU1Sites <- subset(AllSites, AllSites$Lineage == "EU1")

EUV_null = glm(VegBin~1, family = binomial, data=EU1Sites)
EUV_full = glm(VegBin~TrtBin * SoilBin, family=binomial, data=EU1Sites)
step(EUV_null, scope = list(upper=EUV_full), direction="both", test="Chisq", data=EU1Sites)
#VegBin~1

EUV_full2 <- glmulti(VegBin ~ TrtBin * SoilBin, data=EU1Sites, crit=bic, method = "g")
#VegBin~1
EUV_full3 <- glmulti(VegBin ~ TrtBin * SoilBin, data=EU1Sites, crit=aic, method = "g")
#VegBin~1
```

## Logistic Regression for predicting Positive Veg, Generally
```{r}
#Is positive veg predicted by treatment and positive soil &/or interaction? (model from stepwise fit)
LRVeg_TS1 <- glm(VegBin~SoilBin + TrtBin + SoilBin:TrtBin, family=binomial, data=AllSites)
summary(LRVeg_TS1)

#Is positive veg predicted by treatment and positive soil &/or interaction? (model from glmulti fit)
LRVeg_TS2 <- glm(VegBin~1+TrtBin+SoilBin:TrtBin, family=binomial, data=AllSites)
summary(LRVeg_TS2)
```

#Interpretation
Intercept = 1 (as in, yes there's pos veg on plot), it looks like there's a significant predictive quality for whether a site has been treated as to whether there is positive veg. In other words, a treated site has a negative association with positive veg (is less likely to have positive veg than an untreated site). It is plausible that the data eminate from a logistic model that includes these variables. We can reject the null hypothesis that the deviance of the model with these variables is the same as the deviance of the model with only the constant.
> 1-pchisq(79.728-60.026, 58-55)
[1] 0.0001956716

## Logistic regession for predicting positive veg, NA1 only
```{r}
#Is NA1 Positive veg predicted by treatment and positive soil? (model from stepwise fit)
NA1_TS1 <- glm(VegBin~SoilBin + TrtBin, family=binomial, data=NA1Sites)
summary(NA1_TS1)

#Is NA1 Positive veg predicted by treatment and positive soil &/or interaction? (model from glmulti fit)
NA1_TS2 <- glm(VegBin~1+TrtBin+SoilBin:TrtBin, family=binomial, data=NA1Sites)
summary(NA1_TS2)
```

#Interpretation
Intercept = 1 (as in, yes there's pos veg on plot), it looks like there's a significant predictive quality for whether a site has been treated as to whether there is positive veg. In other words, a treated site has a negative association with positive veg (is less likely to have positive veg than an untreated site). It is plausible that the data eminate from a logistic model that includes these variables. We can reject the null hypothesis that the deviance of the model with these variables is the same as the deviance of the model with only the constant.
> 1-pchisq(56.814-41.192, 40-38)
[1] 0.0004052526


## Logistic Regression for predicting positive Veg, EU1 only
```{r}
#Is EU1 Positive veg predicted by any factor?
EU1_TS1 <- glm(VegBin~SoilBin + TrtBin, family=binomial, data=EU1Sites)
summary(EU1_TS1)

summary(EUV_null)
```

#Interpretation
There are no factors with any predictive quality for EU1 positive veg. We accept the null hypothesis that the deviance of the model with these variables is the same as the deviance of the model with only the constant.
> 1-pchisq(19.069-18.421, 17-15)
[1] 0.7232502


# Soil
## Fitting a model for soil
```{r}
soil_null = glm(SoilBin~1, family = binomial, data=AllSites)
soil_full = glm(SoilBin~Lineage * TrtBin * VegBin, family=binomial, data=AllSites)
step(soil_null, scope = list(upper=soil_full), direction="both", test="Chisq", data=AllSItess)
#SoilBin~VegBin

soil_full2 <- glmulti(SoilBin ~ Lineage * TrtBin * VegBin, data=AllSites, crit=bic, method = "g")
#SoilBin~1+VegBin
soil_full3 <- glmulti(SoilBin ~ Lineage * TrtBin * VegBin, data=AllSites, crit=aic, method = "g")
#SoilBin~1+VegBin
```

## Fitting a model for NA1 soil only
```{r}
NAS_null = glm(SoilBin~1, family = binomial, data=NA1Sites)
NAS_full = glm(SoilBin~TrtBin * VegBin, family=binomial, data=NA1Sites)
step(NAS_null, scope = list(upper=NAS_full), direction="both", test="Chisq", data=NA1Sites)
#SoilBin~VegBin

NAS_full2 <- glmulti(SoilBin ~ TrtBin * VegBin, data=NA1Sites, crit=bic, method = "g")
#SoilBin~1+VegBin
NAS_full3 <- glmulti(SoilBin ~ TrtBin * VegBin, data=NA1Sites, crit=aic, method = "g")
#SoilBin~1+VegBin
```

## Fitting a model for EU1 soil only
```{r}
EUS_null = glm(SoilBin~1, family = binomial, data=EU1Sites)
EUS_full = glm(SoilBin~TrtBin * VegBin, family=binomial, data=EU1Sites)
step(EUS_null, scope = list(upper=EUS_full), direction="both", test="Chisq", data=EU1Sites)
#SoilBin ~ 1

EUS_full2 <- glmulti(SoilBin ~ TrtBin * VegBin, data=EU1Sites, crit=bic, method = "g")
#SoilBin ~ 1
EUS_full3 <- glmulti(SoilBin ~ TrtBin * VegBin, data=EU1Sites, crit=aic, method = "g")
#SoilBin ~ 1
```

## Logistic Regression for predicting Positive Soil, Generally
```{r}
#Is positive soil predicted by positive veg alone? (model from stepwise fit)
LRSoil_Veg1 <- glm(SoilBin~VegBin, family=binomial, data=AllSites)
summary(LRSoil_Veg1)

#Is positive soil predicted by positive veg alone? (model from glmulti fit)
LRSoil_Veg2 <- glm(SoilBin~1+VegBin, family=binomial, data=AllSites)
summary(LRSoil_Veg2)
```

#Interpretation
Intercept = 1 (as in, yes there's pos soil on plot), it looks like there's a significant predictive quality for whether a site has positive veg as to whether there is positive soil. In other words, a site with positive veg is more likely to have positive soil than a site without positive veg. It is plausible that the data eminate from a logistic model that includes these variables. We can reject the null hypothesis that the deviance of the model with these variables is the same as the deviance of the model with only the constant.
> 1-pchisq(64.656-54.494, 58-57)
[1] 0.00143365

## Logistic regression for predicting positive soil, NA1 only
```{r}
#Is NA1 Positive soil predicted by positive veg alone? (model from stepwise fit)
NA1_v1 <- glm(SoilBin~VegBin, family=binomial, data=NA1Sites)
summary(NA1_v1)

#Is NA1 Positive soil predicted by positive veg alone? (model from glmulti fit)
NA1_v2 <- glm(SoilBin~1+VegBin, family=binomial, data=NA1Sites)
summary(NA1_v2)
```

#Interpretation
There are no factors with a predictive quality, but at the same time it is plausible that the data eminate from a logistic model that includes these variables. We can reject the null hypothesis that the deviance of the model with these variables is the same as the deviance of the model with only the constant.
> 1-pchisq(37.478-26.734, 40-39)
[1] 0.00104618

## Logistic regression for predicting positive soil, EU1 only
```{r}
EU1_v1 <- glm(SoilBin~VegBin, family=binomial, data=EU1Sites)
summary(EU1_v1)

summary(EUS_null)
```

#Interpretation
There are no factors with any predictive quality for EU1 positive soil. We accept the null hypothesis that the deviance of the model with these variables is the same as the deviance of the model with only the constant.
> 1-pchisq(24.057-23.620, 17-16)
[1] 0.508574