---
title: "Eradication Effectiveness Proportions"
author: "Hazel Daniels"
date: "May 1, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
AllSites <- read_xlsx("BasicData.xlsx")
head(AllSites)
AllSites$VegProp <- as.numeric(AllSites$VegProp)

```

# Vegetation
## Fitting a model for veg
```{r}
veg_null = glm(VegProp~1, family = binomial, data=AllSites)
veg_full = glm(VegProp~Lineage * TrtBin * SoilProp, family=binomial, data=AllSites)
step(veg_null, scope = list(upper=veg_full), direction="both", test="Chisq", data=AllSItess)
#VegProp~SoilProp+TrtBin

library(rJava)
library(glmulti)
model_full2 <- glmulti(VegProp ~ Lineage * TrtBin * SoilProp, data=AllSites, crit=bic, method = "g")
#VegProp~TrtBin+SoilProp
model_full3 <- glmulti(VegProp ~ Lineage * TrtBin * SoilProp, data=AllSites, crit=aic, method = "g")
#VegProp~Lineage+TrtBin+SoilProp+SoilProp:Lineage
```

## Fitting a model for NA1 veg only
```{r}
#Subsetting data
NA1Sites <- subset(AllSites, AllSites$Lineage == "NA1")

NAV_null = glm(VegProp~1, family = binomial, data=NA1Sites)
NAV_full = glm(VegProp~TrtBin * SoilProp, family=binomial, data=NA1Sites)
step(NAV_null, scope = list(upper=NAV_full), direction="both", test="Chisq", data=NA1Sites)
#VegProp~SoilProp

NAV_full2 <- glmulti(VegProp ~ TrtBin * SoilProp, data=NA1Sites, crit=bic, method = "g")
#VegProp~SoilProp
NAV_full3 <- glmulti(VegProp ~ TrtBin * SoilProp, data=NA1Sites, crit=aic, method = "g")
#VegProp~TrtBin+SoilProp
```

## Fitting a model for EU1 veg only
```{r}
#Subsetting data
EU1Sites <- subset(AllSites, AllSites$Lineage == "EU1")

EUV_null = glm(VegProp~1, family = binomial, data=EU1Sites)
EUV_full = glm(VegProp~TrtBin * SoilProp, family=binomial, data=EU1Sites)
step(EUV_null, scope = list(upper=EUV_full), direction="both", test="Chisq", data=EU1Sites)
#VegProp~1

EUV_full2 <- glmulti(VegProp ~ TrtBin * SoilProp, data=EU1Sites, crit=bic, method = "g")
#VegProp~1
EUV_full3 <- glmulti(VegProp ~ TrtBin * SoilProp, data=EU1Sites, crit=aic, method = "g")
#VegProp~1
```

## Logistic Regression for predicting Positive Veg, Generally
```{r}
#Is positive veg predicted by treatment and positive soil &/or interaction? (model from stepwise fit)
LRVeg_TS1 <- glm(VegProp~SoilProp+TrtBin, family=binomial, data=AllSites)
summary(LRVeg_TS1)

#Is positive veg predicted by treatment and positive soil &/or interaction? (model from glmulti fit)
LRVeg_TS2 <- glm(VegProp~Lineage+TrtBin+SoilProp+SoilProp:Lineage, family=binomial, data=AllSites)
summary(LRVeg_TS2)
```

#Interpretation
No significant predictive quality for any factor. 

## Logistic regession for predicting positive veg, NA1 only
```{r}
#Is NA1 Positive veg predicted by treatment and positive soil? (model from stepwise fit)
NA1_TS1 <- glm(VegProp~SoilProp, family=binomial, data=NA1Sites)
summary(NA1_TS1)

#Is NA1 Positive veg predicted by treatment and positive soil &/or interaction? (model from glmulti fit)
NA1_TS2 <- glm(VegProp~TrtBin+SoilProp, family=binomial, data=NA1Sites)
summary(NA1_TS2)
```

#Interpretation (of NA1_TS1)
No significant predictive quality for any factor.

## Logistic Regression for predicting positive Veg, EU1 only
```{r}
#Is EU1 Positive veg predicted by any factor?
EU1_TS1 <- glm(VegProp~SoilProp * TrtBin, family=binomial, data=EU1Sites)
summary(EU1_TS1)

summary(EUV_null)
```

#Interpretation
No significant predictive quality for any factor.

# Soil
## Fitting a model for soil
```{r}
soil_null = glm(SoilProp~1, family = binomial, data=AllSites)
soil_full = glm(SoilProp~Lineage * TrtBin * VegProp, family=binomial, data=AllSites)
step(soil_null, scope = list(upper=soil_full), direction="both", test="Chisq", data=AllSItess)
#SoilProp~1

soil_full2 <- glmulti(SoilProp ~ Lineage * TrtBin * VegProp, data=AllSites, crit=bic, method = "g")
#SoilProp~Lineage+VegProp:Lineage
soil_full3 <- glmulti(SoilProp ~ Lineage * TrtBin * VegProp, data=AllSites, crit=aic, method = "g")
#SoilProp~Lineage+TrtBin+VegProp+VegProp:Lineage+VegProp:TrtBin
```

## Fitting a model for NA1 soil only
```{r}
NAS_null = glm(SoilProp~1, family = binomial, data=NA1Sites)
NAS_full = glm(SoilProp~TrtBin * VegProp, family=binomial, data=NA1Sites)
step(NAS_null, scope = list(upper=NAS_full), direction="both", test="Chisq", data=NA1Sites)
#SoilProp~VegProp

NAS_full2 <- glmulti(SoilProp ~ TrtBin * VegProp, data=NA1Sites, crit=bic, method = "g")
#SoilProp~VegProp
NAS_full3 <- glmulti(SoilProp ~ TrtBin * VegProp, data=NA1Sites, crit=aic, method = "g")
#SoilProp~VegProp
```

## Fitting a model for EU1 soil only
```{r}
EUS_null = glm(SoilProp~1, family = binomial, data=EU1Sites)
EUS_full = glm(SoilProp~TrtBin * VegProp, family=binomial, data=EU1Sites)
step(EUS_null, scope = list(upper=EUS_full), direction="both", test="Chisq", data=EU1Sites)
#SoilProp~1

EUS_full2 <- glmulti(SoilProp ~ TrtBin * VegProp, data=EU1Sites, crit=bic, method = "g")
#SoilProp~1
EUS_full3 <- glmulti(SoilProp ~ TrtBin * VegProp, data=EU1Sites, crit=aic, method = "g")
#SoilProp~1
```

## Logistic Regression for predicting Positive Soil, Generally
```{r}
#Is positive soil predicted by positive veg alone? (model from glmulti fit)
LRSoil_LV1 <- glm(SoilProp~Lineage+VegProp:Lineage, family=binomial, data=AllSites)
summary(LRSoil_LV1)

LRSoil_LV2 <- glm(SoilProp~Lineage+TrtBin+VegProp+VegProp:Lineage+VegProp:TrtBin, family=binomial, data=AllSites)
summary(LRSoil_LV2)

summary(NAS_null)
```

#Interpretation
No significant predictive quality for any factor.

## Logistic regression for predicting positive soil, NA1 only
```{r}
#Is NA1 Positive soil predicted by positive veg alone? (model from stepwise fit & glmulti fit)
NA1_v1 <- glm(SoilProp~VegProp, family=binomial, data=NA1Sites)
summary(NA1_v1)
```

#Interpretation
No significant predictive quality for any factor.

## Logistic regression for predicting positive soil, EU1 only
```{r}
EU1_v1 <- glm(SoilProp~VegProp * TrtBin, family=binomial, data=EU1Sites)
summary(EU1_v1)

summary(EUS_null)
```

#Interpretation
No significant predictive quality for any factor.

#Take Away Message
Binary information is better suited to this analysis.
