---
title: "Curry Co. SOD Analysis"
author: "Hazel Daniels"
date: "14 March 2021"
output:
  word_document: default
  html_document: default
---

```{r setup,include = FALSE}
knitr::opts_chunk$set(echo  =  TRUE)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
#analysis tools
library(readxl)
library(ggplot2)
library(dplyr)
library(vegan)
library(ape)
library(tidyverse)
library(viridis)
library(jtools)
library(car)
library(MASS)
library(gvlma)
library(nlme)
library(broom)
library(rstatix)
library(ggpubr)
library(emmeans)
library(gridExtra)
library(agricolae)
library(scales)

#map-making tools
library(cowplot)
library(googleway)
library(ggrepel)
library(ggspatial)
library(lwgeom)
library(maps)
library(sf)
library(rgdal)
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)
library(tools)
```

# Reading and manipulating data
```{r,message = FALSE,warning = FALSE}
SODSites <- read_xlsx("BasicData.xlsx",1)
SODSite.ord <- SODSites[-c(31),,drop = F] 
#removing HD-4331 as it skews results strongly

# Subsetting data
AllSites <- SODSites[-c(74,76,77,80,81,82,86,95,96,97),,drop = F] 
#removing HD-4376, 4378, 4379, 4382, 4383, 4384, 4388, 4397, 4398, 4399 as these sites were measured before and after treatment (removing treated data). Removing these duplicate sites to a data subset for separate analysis.
AllSites$TrtBin <- as.factor(AllSites$TrtBin)

NA1 <- subset(AllSites, AllSites$Lineage == "NA1")

AllSeeds <- read_xlsx("BasicData.xlsx",2)
AllSeeds <- AllSeeds[-13] 
#removing CE as it skews results strongly
AllSeeds <- AllSeeds[-c(31),,drop = F] 
#removing HD-4331 as it skews results strongly
AllSeeds$TrtBin <- as.factor(AllSeeds$TrtBin)
AllSeeds$WFBin <- as.factor(AllSeeds$WFBin)
AllSeeds$VegBin <- as.factor(AllSeeds$VegBin)
AllSeeds$SoilBin <- as.factor(AllSeeds$SoilBin)

AllFire <- read_xlsx("BasicData.xlsx",3)
AllFire$RX <- as.factor(AllFire$RX)

EU1wide <- read_xlsx("BasicData.xlsx",4)

EU1long <- read_xlsx("BasicData.xlsx",5)
```

# Visualization by Analysis Group
```{r}
OCW <- st_as_sf(map("state", plot = FALSE, fill = TRUE, 
                    crs = 4326, agr = "constant"))
OCW <- OCW[c(4,36,46),,]
OCW$area <- as.numeric(st_area(OCW))
OCW$ID <- toTitleCase(OCW$ID)
OCW$X <- c(-124.1,-124.1, -124.1)
OCW$Y <- c(41.99, 43, 44)
OCW$X2 <- c(-123,-123, -122.5)
OCW$Y2 <- c(41, 44, 46.5)

LatLong <- SODSites %>% select(`New Site`, 
                               Lineage, 
                               TrtBin, 
                               Latitude, 
                               Longitude)
LatLong$Grp <- c(1,2,2,2,2,2,2,2,2,2,2,2,2,2,2,2,3,3,2,1,3,3,3,2,2,2,2,3,1,1,2,2,2,2,3,3,1,2,1,2,2,3,2,1,2,2,2,2,2,1,2,1,2,3,2,2,2,2,2,2,2,2,2,2,2,2,2,2,2,2,2,1,1,3,1,3,3,1,1,3,3,3,1,1,1,3,1,1,1,1,1,1,1,1,3,3,3,1)
LatLong$Latitude <- as.numeric(LatLong$Latitude)
LatLong$Longitude <- as.numeric(LatLong$Longitude)
LatLong$Longitude <- LatLong$Longitude + 0.05 
#forced offset since the layers weren't playing nice :(
LatLong <- st_as_sf(LatLong, coords = c("Longitude", "Latitude"),
                    crs = 4326, agr = "constant")


ORci <- data.frame(state = rep("Oregon", 2), 
                   city = c("Brookings", "Gold Beach"), 
                   lat = c(42.0526, 42.4073), 
                   lng = c(-124.22, -124.37))
ORci <- st_as_sf(ORci, coords = c("lng", "lat"), remove = FALSE, 
    crs = 4326, agr = "constant")

inset <- ggplot(data = OCW)+
    geom_sf(fill = "cornsilk")+
    geom_text(data = OCW, aes(X2, Y2, label = ID), size = 3)+
    coord_sf(xlim = c(-124.5, -121.5), ylim = c(41, 47))+
    geom_rect(xmin = -124.5, xmax = -124, 
              ymin = 42, ymax = 42.5,
              fill = NA, colour = "black", size = 1) +
    theme(panel.grid.major = element_line(color = gray(0.5), 
          linetype = "dashed",size = 0.5), 
          panel.background = element_rect(fill = "aliceblue"),
          axis.ticks = element_blank(),
          axis.text.x = element_blank(), 
          axis.text.y = element_blank(),
          axis.title.x = element_blank(),
          axis.title.y = element_blank())

map <- ggplot(data = OCW)+
    geom_sf(fill = "antiquewhite1")+
    geom_sf(data = LatLong, size = 3, shape = 22, aes(fill=factor(Grp)))+
    geom_sf(data = ORci)+
    geom_text(data = OCW, aes(X, Y, label = ID), size = 4)+
    geom_text_repel(data = ORci, 
                    aes(x = lng, y = lat, label = city), 
                    fontface = "bold", 
                    nudge_x = c(-0.05,0.05), nudge_y = c(-0.01,-0.01))+
    scale_fill_viridis_d(option = "plasma", 
                         name = "Analysis Groups", 
                         breaks = c("1", "3", "2"), 
                         labels = c("EU1", "EU1 Resample", "NA1"))+
    annotation_scale(location = "bl", width_hint = 0.4)+
    annotation_north_arrow(location = "tl", which_north = "true", 
        pad_x = unit(0.3, "in"), pad_y = unit(0.35, "in"),
        style = north_arrow_fancy_orienteering)+
    coord_sf(xlim = c(-124.5, -124), ylim = c(42, 42.5))+
    xlab("Longitude")+ylab("Latitude")+
    theme(panel.grid.major = element_line(color = gray(0.5), 
          linetype = "dashed",size = 0.5), 
          panel.background = element_rect(fill = "aliceblue"),
          legend.position = c(0.17,0.17))

mapinset <- ggdraw(map) +
    draw_plot(inset, width = 0.45, height = 0.45, x = 0.565, y = 0.53)

ggsave(filename = "MapInset.png",
       mapinset,
       width = 6.5,
       height = 6.5,
       dpi = 300,
       units = "in",
       device = 'png')
```

# EU1 NA1 Mixed Group
## Visualizing Data
```{r,echo = FALSE,message = FALSE,warning = FALSE}
dv <- AllSites %>% select(Lineage,VegProp,TrtBin)

dvsum <- dv %>% 
  group_by(Lineage,TrtBin) %>%
  summarise(n = n(),mean = mean(VegProp),sd = sd(VegProp)) %>%
  mutate(se = sd/sqrt(n))  %>%
  mutate(ic = se*qt((1-0.05)/2+0.5, n-1))

pV <- ggplot(dvsum,aes(x = TrtBin,
                       y = mean,
                       fill = Lineage,
                       ymin = mean-se,
                       ymax = mean+se))+
  geom_bar(stat = "identity",position = "dodge",color = "gray",width = 0.9)+
  ggtitle("A.")+
  scale_fill_viridis_d(name = "Lineage", guide = FALSE)+
  scale_x_discrete(name = "Treatment",
                   breaks = c("0","1"), 
                   labels = c("0" = "Untreated","1" = "Treated"))+
  scale_y_continuous(labels = scales::number_format
                     (accuracy = 0.01),
                     limits = c(0,0.6))+
  theme_classic()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_blank(),
        strip.text.x  =  element_blank())+
  facet_wrap(~Lineage,ncol = 2)
  
pVV <- pV+geom_errorbar(width = 0.2,colour = "orange",alpha = 0.9,size = 1.3)

ds <- AllSites %>% select(Lineage,SoilProp,TrtBin)

dssum <- ds %>% 
  group_by(Lineage,TrtBin) %>%
  summarise(n = n(),mean = mean(SoilProp),sd = sd(SoilProp)) %>%
  mutate(se = sd/sqrt(n))  %>%
  mutate(ic = se*qt((1-0.05)/2 +0.5,n-1))

pS <- ggplot(dssum,aes(x = TrtBin,
                       y = mean,
                       fill = Lineage,
                       ymin = mean-se,
                       ymax = mean+se))+
  geom_bar(stat = "identity",position = "dodge",color = "gray",width = 0.9)+
  ylim(0,0.06)+
  ggtitle("B.")+
  labs(x = "Treatment")+
  scale_fill_viridis_d(name = "Lineage")+
  scale_x_discrete(name = "Treatment",
                   breaks = c("0","1"),
                   labels = c("0" = "Untreated","1" = "Treated"))+
  theme_classic()+ 
  theme(legend.position = "bottom",
        legend.title = element_blank(),
        axis.title.y = element_blank(),
        strip.background = element_blank(),
        strip.text.x  =  element_blank())+
  facet_wrap(~Lineage,ncol = 2)

pSS <- pS+geom_errorbar(width = 0.2,colour = "orange",alpha = 0.9,size = 1.3)

pVS <- ggarrange(pVV,pSS,ncol = 1,nrow = 2,
                 heights = c(0.7,1))
plotVS<- annotate_figure(pVS,left = text_grob("Mean proportion of source material",rot = 90))
plotVS

ggsave(filename = "MeanProp.png",
       plotVS,
       width = 6.5,
       height = 6,
       dpi = 300,
       units = "in",
       device = 'png')
```

## Veg Mix
```{r,message = FALSE,warning = FALSE}
V1 <- lm(VegProp~Lineage*TrtBin,data = AllSites)

anova(V1)
summary(V1)
summ(V1)

# Least Squares Means
V1.lsm <- lsmeans(V1,"TrtBin")
V1LSM <- as.data.frame(V1.lsm)

#Least Squares Means visualization
pLSV <- ggplot(V1LSM,aes(x = TrtBin,
                         y = lsmean,
                         fill = TrtBin,
                         ymin = lower.CL,
                         ymax = upper.CL))+
  geom_bar(stat = "identity",position = "dodge",color = "gray",width = 0.9)+
  ylim(0,0.6)+
  ggtitle("Vegetation")+
  scale_fill_viridis_d(name = "Lineage",guide = FALSE,)+
  scale_x_discrete(name = "Treatment",
                   breaks = c("0","1"),
                   labels = c("0" = "Untreated","1" = "Treated"))+
  theme_classic()+
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.y = element_blank())
  
pLSVV <- pLSV+geom_errorbar(width = 0.2,colour = "orange",alpha = 0.9,size = 1.3)

#Contrasts
V1.c <- contrast(V1.lsm,list(TvU = c(-1,1)))
V1.c

#correlations of random effects not included in the model
VC1 <- cor.test(AllSites$VegProp,AllSites$YrSinceTrt, method = c("pearson","kendall","spearman"))
VC11 <- cor.test(AllSites$VegProp,AllSites$TotalTO, method = c("pearson","kendall","spearman"))
VC111 <- cor.test(AllSites$VegProp,AllSites$Num_Seedlings, method = c("pearson","kendall","spearman"))
VC1111 <- cor.test(AllSites$VegProp,AllSites$SoilBin, method = c("pearson","kendall","spearman"))
VC1
VC11
VC111
VC1111
#correlation between soil and veg!
```

## Soil Mix
```{r,message = FALSE,warning = FALSE}
S1 <- glm(SoilBin~Lineage*TrtBin,family = "binomial",data = AllSites)

anova(S1)
summary(S1)
summ(S1)

# Contrast statement
S1.lsm <- lsmeans(S1,"TrtBin")
S1.lsm
S1LSM <- as.data.frame(S1.lsm)
S1LSM

#Least Squares Means visualization
pLSS <- ggplot(S1LSM,aes(x = TrtBin,
                         y = lsmean,
                         fill = TrtBin,
                         ymin = asymp.LCL,
                         ymax = asymp.UCL))+
  geom_bar(stat = "identity",position = "dodge",color = "gray",width = 0.9)+
  ylim(-2.5,0)+
  ggtitle("Soil")+
  scale_fill_viridis_d(name = "Lineage",guide = FALSE,)+
  scale_x_discrete(name = "Treatment",
                   breaks = c("0","1"),
                   labels = c("0" = "Untreated","1" = "Treated"))+
  theme_classic()+
  theme(axis.title.y = element_blank())
  
pLSSS <- pLSS+geom_errorbar(width = 0.2,colour = "orange",alpha = 0.9,size = 1.3)

#Combining veg and soil plots
pLSVS <- ggarrange(pLSVV,pLSSS,ncol = 1,nrow = 2)
plotLSVS<- annotate_figure(pLSVS,left = text_grob("Least Squares Mean",rot = 90))
plotLSVS

ggsave(filename = "LSM.png",
       plotLSVS,
       width = 6.5,
       height = 5,
       dpi = 300,
       units = "in",
       device = 'png')

#Contrasts
S1.c <- contrast(S1.lsm,list(TvU = c(-1,1)))
S1.c

#correlations of random effects not included in the model
SC1 <- cor.test(AllSites$SoilBin,AllSites$YrSinceTrt, method = c("pearson","kendall","spearman"))
SC11 <- cor.test(AllSites$SoilBin,AllSites$TotalTO, method = c("pearson","kendall","spearman"))
SC111 <- cor.test(AllSites$SoilBin,AllSites$Num_Seedlings, method = c("pearson","kendall","spearman"))
SC1
SC11
SC111
```

## V Mix Assumptions
```{r,message = FALSE,warning = FALSE}
plot(V1)
gvlma(V1)

#linearity
ggplot(AllSites,aes(TrtBin,VegProp))+ 
  stat_smooth(method = "loess")+ 
  stat_smooth(method = "lm",color = "red",fill = "red",alpha = .25)+ 
  geom_point()

#constant variance
spreadLevelPlot(V1)

#normality of residuals
V1.resi <- qqPlot(V1,main = "QQ Plot")

V1.res <- studres(V1)
hist(V1.res,freq = FALSE,main = "Distribution of Studentized Residuals")
V1.xfit <- seq(min(V1.res),max(V1.res),length  =  40)
V1.yfit <- dnorm(V1.xfit)
lines(V1.xfit,V1.yfit)

#multicollinearity
vif(V1)
sqrt(vif(V1))>2

#outliers
AllSites$Vleverage <- hatvalues(V1)
ggplot(AllSites,aes(Plot,Vleverage))+geom_point()+ylim(0,1)+xlab("test")
```

## S Mix Assumptions
```{r,message = FALSE,warning = FALSE}
plot(S1)

#linearity
ggplot(AllSites,aes(SoilBin,TrtBin))+ 
  stat_smooth(method = "loess")+ 
  stat_smooth(method = "lm",color = "red",fill = "red",alpha = .25)+ 
  geom_point()

#constant varience
spreadLevelPlot(S1)

#normality of residuals
S1.resi <- studres(S1)
hist(S1.resi,freq  =  FALSE,main = "Distribution of Studentized Residuals")
S1.xfit <- seq(min(S1.resi),max(S1.resi),length = 40)
S1.yfit <- dnorm(S1.xfit)
lines(S1.xfit,S1.yfit)
#assumption violated?

#multicollinearity
vif(S1)
sqrt(vif(S1))>2

#outliers
AllSites$Sleverage <- hatvalues(S1)
ggplot(AllSites,aes(Plot,Sleverage))+geom_point()+ylim(0,1)+xlab("test")
```

# EU1 Repeated Measures
## EU1 Veg
```{r,message = FALSE,warning = FALSE}
V2 <- lme(VegProp~TrtBin,random = ~1|Plot,data = EU1long)

anova(V2)
summary(V2)

V.dt <-array(c(1,0,8,1),
           dim = c(2,2),
           dimnames = list(
             Time1 = c("FirstPos","FirstNeg"),
             Time2 = c("ThenPos","ThenNeg")))
V.dt

mcnemar.test(V.dt,y = NULL,correct = FALSE)
```

## EU1 Soil
```{r,message = FALSE,warning = FALSE}
S2 <- lme(SoilBin~TrtBin,random = ~1|Plot,data = EU1long)

anova(S2)
summary(S2)

S.dt <-array(c(1,1,4,4),
           dim = c(2,2),
           dimnames = list(
             Time1 = c("FirstPos","FirstNeg"),
             Time2 = c("ThenPos","ThenNeg")))
S.dt

mcnemar.test(S.dt,y = NULL,correct = FALSE)
```

# NA1 Wildfire
## NA1 Veg
```{r,message = FALSE,warning = FALSE}
V3 <- lm(VegProp~RX,data = AllFire)

anova(V3)
summary(V3)
summ(V3)

# Contrast statement
levels(AllFire$RX)
# 0 - No Wildfire or Treatment
# 1 - Only Treatment
# 2 - Only Wildfire
# 3 - Both Wildfire and Treatment

#Least Squares Means
V3.lsm <- lsmeans(V3,~RX,adjst = "tukey")
V3.lsm
V3LSM <- as.data.frame(V3.lsm)
V3LSM

#Contrasts
V3.c <- contrast(V3.lsm,list(
  "0 v. 1" = c(1,-1,0,0), #difference is RX
  "0 v. 2" = c(1,0,-1,0), #difference is WF
  "0 v. 3" = c(1,0,0,-1), #difference is both
  "1 v. 2" = c(0,-1,1,0), #difference is both
  "1 v. 3" = c(0,1,0,-1), #difference is WF
  "2 v. 3" = c(0,0,1,-1), #difference is RX
  "1&3 v. 0" = c(1,-0.5,0,-0.5), #both RX to noWF or RX
  "2&3 v. 0" = c(1,0,-0.5,-0.5))) #both WF to noWF or RX
V3.c

V3CON <- as.data.frame(V3.c)
V3CON$contrast <- factor(V3CON$contrast, levels=c("0 v. 1","0 v. 2","0 v. 3","1 v. 2","1 v. 3","2 v. 3","1&3 v. 0","2&3 v. 0"))

p.VWF <- ggplot(V3CON,aes(x = contrast,
                 y = estimate,
                 fill = contrast,
                 ymin = estimate-SE,
                 ymax = estimate+SE))+
  geom_bar(stat = "identity",position = "dodge",color = "gray")+
  scale_fill_viridis_d()+theme_classic()+
  scale_x_discrete(name = "Contrast Statement",
                   guide = guide_axis(angle = 20),
                   labels = c(
                     "0 v. 1" = "None v. RX",
                     "0 v. 2" = "None v. WF",
                     "0 v. 3" = "None v. Both",
                     "1 v. 2" = "RX v. WF",
                     "1 v. 3" = "RX v. Both",
                     "2 v. 3" = "WF v. Both",
                     "1&3 v. 0" = "RX + Both v. None",
                     "2&3 v. 0" = "WF + Both v. None"))+
  scale_y_continuous("Least Squares Mean",
                     limits = c(-0.1,0.55),
                     breaks = c(-0.1,0,0.1,0.2,0.3,0.4,0.5))+
  theme(legend.position = "none",
        text = element_text(size = 15))+
  geom_errorbar(width = 0.2,colour = "orange",alpha = 0.9,size = 1.3)+
  geom_text(aes(x = contrast,y = estimate+SE+0.01,
                label = c("*","","*","","","","*","*")),
            size = 5,position = position_dodge(.5))
p.VWF

ggsave(filename = "WFVeg.png",
       p.VWF,
       width = 6.5,
       height = 5,
       dpi = 300,
       units = "in",
       device = 'png')

#Correlations of random effects not included in the model
VC3 <- cor.test(AllFire$VegProp,AllFire$TotalTO,method = c("pearson","kendall","spearman"))
VC33 <- cor.test(AllFire$VegProp,AllFire$SoilProp,method = c("pearson","kendall","spearman"))
VC3
VC33 #correlation between soil and veg!
```

## NA1 Soil
```{r,message = FALSE,warning = FALSE}
S3 <- glm(SoilBin~RX,data = AllFire)

anova(S3)
summary(S3)
summ(S3)

# Contrast statement
S3.lsm <- lsmeans(S3,~RX,adjst = "tukey")
S3.lsm
S3LSM <- as.data.frame(S3.lsm)
S3LSM

S3.c <- contrast(S3.lsm,list(
  "0 v. 1" = c(-1,1,0,0), #difference is RX
  "0 v. 2" = c(-1,0,1,0), #difference is WF
  "0 v. 3" = c(1,0,0,-1), #difference is both
  "1 v. 2" = c(0,-1,1,0), #difference is both
  "1 v. 3" = c(0,1,0,-1), #difference is WF
  "2 v. 3" = c(0,0,1,-1), #difference is RX
  "1&3 v. 0" = c(1,-0.5,0,-0.5), #both RX to noWF or RX
  "2&3 v. 0" = c(-1,0,0.5,0.5))) #both WF to noWF or RX
S3.c

S3CON <- as.data.frame(S3.c)
S3CON$contrast <- factor(S3CON$contrast, levels=c("0 v. 1","0 v. 2","0 v. 3","1 v. 2","1 v. 3","2 v. 3","1&3 v. 0","2&3 v. 0"))

p.SWF <- ggplot(S3CON,aes(x = contrast,
                 y = estimate,
                 fill = contrast,
                 ymin = estimate-SE,
                 ymax = estimate+SE))+
  geom_bar(stat = "identity",position = "dodge",color = "gray")+
  scale_fill_viridis_d()+theme_classic()+
  scale_x_discrete(name = "Contrast Statement",
                   guide = guide_axis(angle = 20),
                   labels = c(
                     "0 v. 1" = "None v. RX",
                     "0 v. 2" = "None v. WF",
                     "0 v. 3" = "None v. Both",
                     "1 v. 2" = "RX v. WF",
                     "1 v. 3" = "RX v. Both",
                     "2 v. 3" = "WF v. Both",
                     "1&3 v. 0" = "RX + Both v. None",
                     "2&3 v. 0" = "WF + Both v. None"))+
  scale_y_continuous("Least Squares Mean",
                     limits = c(-0.11,0.8),
                     breaks = c(-0.1,0,0.2,0.4,0.6,0.8))+
  theme(legend.position = "none",
        text = element_text(size = 15))+
  geom_errorbar(width = 0.2,colour = "orange",alpha = 0.9,size = 1.3)+
  geom_text(aes(x = contrast,y = estimate+SE+0.01,
                label = c("","*","","*","","*","","")),
            size = 5,position = position_dodge(.5))

p.SWF

ggsave(filename = "WFSoil.png",
       p.SWF,
       width = 6.5,
       height = 5,
       dpi = 300,
       units = "in",
       device = 'png')

#correlations of random effects not included in the model
SC3 <- cor.test(AllFire$VegProp,AllFire$TotalTO,method = c("pearson","kendall","spearman"))
SC33 <- cor.test(AllFire$VegBin,AllFire$SoilBin,method = c("pearson","kendall","spearman"))
SC3
SC33 #correlation between soil and veg!
```

## NA1 V Assumptions
```{r,message = FALSE,warning = FALSE}
plot(V3)
gvlma(V3)

#constant variance
spreadLevelPlot(V3)

#normality of residuals
V3.resi <- qqPlot(V3, main = "QQ Plot")

V3.res <- studres(V3)
hist(V3.res, freq = FALSE,
     main = "Distribution of Studentized Residuals")
V3.xfit <- seq(min(V3.res), max(V3.res), length  =  40)
V3.yfit <- dnorm(V3.xfit)
lines(V3.xfit,V3.yfit)

#outliers
AllFire$WVleverage <- hatvalues(V3)
ggplot(AllFire,aes(`New Site`,WVleverage))+
  geom_point()+ylim(0,1)+xlab("test")
```

## NA1 S Assumptions
```{r,message = FALSE,warning = FALSE}
plot(S3)

#constant variance
spreadLevelPlot(S3)

#normality of residuals
S3.res <- studres(S3)
hist(S3.res,freq = FALSE,
     main = "Distribution of Studentized Residuals")
S3.xfit <- seq(min(S3.res),max(S3.res),length = 40)
S3.yfit <- dnorm(S3.xfit)
lines(S3.xfit,S3.yfit)

#outliers
AllFire$WSleverage <- hatvalues(S3)
ggplot(AllFire,aes(`New Site`,WSleverage))+
  geom_point()+ylim(0,1)+xlab("test")
```

# Seedling Diversity
## Alpha Diversity
```{r,message = FALSE,warning = FALSE}
PCASeeds <- AllSeeds[-(2:6)]
PCASeeds <- column_to_rownames(PCASeeds, var  =  "New Site")

#different alpha diversity statistics
AllSeeds$Shannon <- diversity(PCASeeds,index = "shannon")

hist(AllSeeds$Shannon)
ggplot(AllSeeds,aes(x = TrtBin,y = Shannon))+geom_boxplot()

#TrtBin
#anova
ShaTrt.a <- aov(Shannon~TrtBin,AllSeeds)
summary.lm(ShaTrt.a)
#Tukey
ShaTrt.t <- HSD.test(ShaTrt.a,"TrtBin")
print(ShaTrt.t)

#Lineage
#anova
ShaLin.a <- aov(Shannon~Lineage,AllSeeds)
summary.lm(ShaLin.a)
#Tukey
ShaLin.t <- HSD.test(ShaLin.a,"Lineage")
print(ShaLin.t)

#Site
#anova
ShaSite.a <- aov(Shannon~`New Site`,AllSeeds)
summary(ShaSite.a)
#Tukey
#not possible as there are 0 df remaining

compare_alpha <- function(AllSeeds,grouping_var) {
  AllSeeds$Shannon <- diversity(PCASeeds,index = "shannon")
  AllSeeds$grouping <- AllSeeds[[grouping_var]]
  anova_result <- aov(Shannon~grouping,AllSeeds)
  tukey_result <- HSD.test(anova_result,"grouping",group = TRUE)
  group_data <- tukey_result$groups[order(rownames(tukey_result$groups)),]
  my_plot <- ggplot(AllSeeds,aes(x = grouping,y = Shannon))+
    geom_text(data = data.frame(),
              aes(x = rownames(group_data),
                  y = max(AllSeeds$Shannon)+1,
                  label = group_data$groups),
              col = 'black',
              size = 10)+
    geom_boxplot()+
    ggtitle("Alpha diversity")+
    xlab(grouping_var)+
    ylab("Alpha diversity index (Shannon)")
  return(my_plot)
}

compare_alpha(AllSeeds,"Lineage")
compare_alpha(AllSeeds,"TrtBin")
compare_alpha(AllSeeds,"WFBin")
compare_alpha(AllSeeds,"VegBin")
compare_alpha(AllSeeds,"SoilBin")
```

## Beta Diversity
### PCA
```{r,message = FALSE,warning = FALSE}
PCA <- rda(PCASeeds,scale = FALSE)
barplot(as.vector(PCA$CA$eig)/sum(PCA$CA$eig)) 
sum((as.vector(PCA$CA$eig)/sum(PCA$CA$eig))[1:2]) # 82%

sitePCA <- PCA$CA$u # Site scores
speciesPCA <- PCA$CA$v # Species scores

biplot(PCA,choices = c(1,2),type = c("text","points"),xlim = c(-5,10)) 
# biplot of axis 1 vs 2

biplot(PCA,choices = c(1,3),type = c("text","points")) 
# biplot of axis 1 vs 3
```

### PCoA
```{r,message = FALSE,warning = FALSE}
DistSeed <- vegdist(PCASeeds,method = "bray")
PCOA <- pcoa(DistSeed)

barplot(PCOA$values$Relative_eig[1:10])
sum((as.vector(PCOA$values$Relative_eig)/sum(PCOA$values$Relative_eig))[1:5]) #63%

biplot.pcoa(PCOA,PCASeeds)
```

### NMDS
```{r,message = FALSE,warning = FALSE}
NMDS <- metaMDS(PCASeeds,
                k = 2,
                trymax = 100,
                trace = F,
                autotransform = FALSE,
                distance = "bray")

EFvars <- c("New Site",
            "Lineage",
            "Latitude",
            "Longitude",
            "VegBin",
            "SoilBin",
            "YrSinceTrt",
            "TrtBin")

EFSites <- SODSite.ord[EFvars]

ef <- envfit(NMDS,EFSites,permutations = 999,na.rm = TRUE)

data.scores <- as.data.frame(scores(NMDS))
data.scores$Site <- rownames(data.scores)
data.scores$Treatment <- AllSeeds$TrtBin
data.scores$Lineage <- AllSeeds$Lineage
data.scores$VegBin <- AllSeeds$VegBin
data.scores$SoilBin <- AllSeeds$SoilBin

species.scores <- as.data.frame(scores(NMDS, "species"))
species.scores$Species <- rownames(species.scores)

nmds.un <- data.scores[data.scores$Treatment == "0",
                       ][chull(data.scores
                               [data.scores$Treatment == "0", 
                   c("NMDS1", "NMDS2")]), ]
nmds.tr <- data.scores[data.scores$Treatment == "1",
                       ][chull(data.scores
                               [data.scores$Treatment == "1", 
                   c("NMDS1", "NMDS2")]), ]
hull.data <- rbind(nmds.un, nmds.tr)

p.nmds <- ggplot()+ 
  geom_polygon(data = hull.data,
               aes(x = NMDS1, y = NMDS2,
                   fill = Treatment,
                   group = Treatment),
               alpha = 0.3)+
  geom_point(data = data.scores,
             aes(x = NMDS1, y=NMDS2,
                 shape = as.factor(Lineage),
                 color = Treatment),
             size=4)+
  geom_text(data = species.scores,
            aes(x = NMDS1, y = NMDS2,
                label = Species),
            alpha = 0.5)+
  coord_equal()+
  labs(shape = "Lineage")+
  scale_fill_viridis(discrete = TRUE, labels = c("0" = "Untreated","1" = "Treated"))+
  scale_color_viridis(discrete = TRUE, begin = 0.3, end = 0.8, labels = c("0" = "Untreated","1" = "Treated"))+
  theme_classic()+
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title.x = element_text(size=18),
        axis.title.y = element_text(size=18),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.background = element_blank())
p.nmds

ggsave(filename = "NMDS.png",
       p.nmds,
       width = 6.5,
       height = 5,
       dpi = 300,
       units = "in",
       device = 'png')
```

### PERMANOVA
```{r,message = FALSE,warning = FALSE}
PER.div <- adonis2(PCASeeds~TrtBin*Lineage,data = EFSites, 
                   permutations = 999,method = "bray")
PER.div

# overall tests
PER2.div <- adonis2(PCASeeds~TrtBin*Lineage,data = EFSites,
                    permutations = 999,method = "bray",by = NULL)
PER2.div
```

#### Homoscedasticity test by Treatment
```{r}
T.disp <- betadisper(DistSeed, group = EFSites$TrtBin)
permutest(T.disp)
plot(T.disp, hull = FALSE, ellipse = TRUE) #sd ellipse

T.cent <- data.frame(grps = rownames(T.disp$centroids),
                        data.frame(T.disp$centroids))
T.vec <- data.frame(group = T.disp$group,
                      data.frame(T.disp$vectors))
T.vec$Lineage <- AllSeeds$Lineage

seg.un <- T.vec[T.vec$group == "0",]
seg.tr <- T.vec[T.vec$group == "1",]
seg.u <- cbind(seg.un[,1:3],
                  T.cent[rep(1:nrow(T.cent), 
                  as.data.frame(table(seg.un$group))$Freq),2:3])
seg.t <- cbind(seg.tr[,1:3],
                  T.cent[rep(1:nrow(T.cent), 
                  as.data.frame(table(seg.tr$group))$Freq),2:3])
names(seg.u)<-c("group","v.PCoA1","v.PCoA2","PCoA1","PCoA2")
names(seg.t)<-c("group","v.PCoA1","v.PCoA2","PCoA1","PCoA2")
seg.data <- rbind(seg.u, seg.t)

un.hull<-seg.data[seg.data$group=="0",1:3][chull(seg.data[seg.data$group=="0",2:3]),]
tr.hull<-seg.data[seg.data$group=="1",1:3][chull(seg.data[seg.data$group=="1",2:3]),]
trt.hull<-rbind(un.hull,tr.hull)

panel.un <- ggplot()+ 
  geom_polygon(data=trt.hull[trt.hull=="0",],
               aes(x=v.PCoA1,y=v.PCoA2),
               colour="black",alpha=0,linetype="dashed")+
  geom_segment(data=seg.data[seg.data$group == "0",],
               aes(x=v.PCoA1,xend=PCoA1,y=v.PCoA2,yend=PCoA2),
               alpha=0.30)+ 
  geom_point(data=T.cent[1,1:3],
             aes(x=PCoA1,y=PCoA2),
             size=4,colour="red",shape=16)+
  geom_point(data=seg.data[seg.data$group == "0",],
             aes(x=v.PCoA1,y=v.PCoA2),size=2,shape=16)+
  labs(title="A.",x="",y="")+
  coord_cartesian(xlim = c(-0.5,0.5), ylim = c(-0.5,0.5))+
  theme_bw()+ 
  theme(legend.position="none")
panel.tr <- ggplot()+ 
  geom_polygon(data=trt.hull[trt.hull=="1",],
               aes(x=v.PCoA1,y=v.PCoA2),
               colour="black",alpha=0,linetype="dashed")+
  geom_segment(data=seg.data[seg.data$group == "1",],
               aes(x=v.PCoA1,xend=PCoA1,y=v.PCoA2,yend=PCoA2),
               alpha=0.30)+ 
  geom_point(data=T.cent[2,1:3],
             aes(x=PCoA1,y=PCoA2),
             size=4,colour="red",shape=17)+ 
  geom_point(data=seg.data[seg.data$group == "1",],
             aes(x=v.PCoA1,y=v.PCoA2),size=2,shape=17)+
  labs(title="B.",x="",y="")+
  coord_cartesian(xlim = c(-0.5,0.5), ylim = c(-0.5,0.5))+
  theme_bw()+ 
  theme(legend.position="none")
panel.all <- ggplot()+ 
  geom_polygon(data=trt.hull,
               aes(x=v.PCoA1,y=v.PCoA2),
               colour="black",alpha=0,linetype="dashed")+
  geom_segment(data=seg.data,
               aes(x=v.PCoA1,xend=PCoA1,y=v.PCoA2,yend=PCoA2),
               alpha=0.30)+ 
  geom_point(data=T.cent[,1:3],
             aes(x=PCoA1,y=PCoA2,shape=grps),
             size=4,colour="red")+ 
  geom_point(data=seg.data,
             aes(x=v.PCoA1,y=v.PCoA2,shape=group),size=2)+
  labs(title="C.",x="",y="")+
  coord_cartesian(xlim = c(-0.5,0.5), ylim = c(-0.5,0.5))+
  theme_bw()+ 
  theme(legend.position="none")

p.perT <- ggarrange(panel.un,panel.tr,panel.all,ncol = 3,nrow = 1)
p.perT

ggsave(filename = "Perm_Trt.png",
       p.perT,
       width = 6.5,
       height = 2,
       dpi = 300,
       units = "in",
       device = 'png')
```

#### By Lineage
```{r}
L.disp <- betadisper(DistSeed, group = EFSites$Lineage)
permutest(L.disp)
plot(L.disp, hull = FALSE, ellipse = TRUE) #sd ellipse

L.cent <- data.frame(grps = rownames(L.disp$centroids),
                        data.frame(L.disp$centroids))
L.vec <- data.frame(group = L.disp$group,
                      data.frame(L.disp$vectors))
L.vec$Lineage <- AllSeeds$Lineage

seg.eu <- L.vec[L.vec$group == "EU1",]
seg.na <- L.vec[L.vec$group == "NA1",]
seg.eu <- cbind(seg.eu[,1:3],
                  L.cent[rep(1:nrow(L.cent), 
                  as.data.frame(table(seg.eu$group))$Freq),2:3])
seg.na <- cbind(seg.na[,1:3],
                  L.cent[rep(1:nrow(L.cent), 
                  as.data.frame(table(seg.na$group))$Freq),2:3])
names(seg.eu)<-c("group","v.PCoA1","v.PCoA2","PCoA1","PCoA2")
names(seg.na)<-c("group","v.PCoA1","v.PCoA2","PCoA1","PCoA2")
lin.data <- rbind(seg.eu, seg.na)

eu.hull<-lin.data[lin.data$group=="EU1",1:3][chull(lin.data[lin.data$group=="EU1",2:3]),]
na.hull<-lin.data[lin.data$group=="NA1",1:3][chull(lin.data[lin.data$group=="NA1",2:3]),]
lin.hull<-rbind(eu.hull,na.hull)

panel.eu <- ggplot()+ 
  geom_polygon(data=lin.hull[lin.hull=="EU1",],
               aes(x=v.PCoA1,y=v.PCoA2),
               colour="black",alpha=0,linetype="dashed")+
  geom_segment(data=lin.data[lin.data$group == "EU1",],
               aes(x=v.PCoA1,xend=PCoA1,y=v.PCoA2,yend=PCoA2),
               alpha=0.30)+ 
  geom_point(data=L.cent[1,1:3],
             aes(x=PCoA1,y=PCoA2),
             size=4,colour="red",shape=16)+
  geom_point(data=lin.data[lin.data$group == "EU1",],
             aes(x=v.PCoA1,y=v.PCoA2),size=2,shape=16)+
  labs(title="A.",x="",y="")+
  coord_cartesian(xlim = c(-0.5,0.5), ylim = c(-0.5,0.5))+
  theme_bw()+ 
  theme(legend.position="none")
panel.na <- ggplot()+ 
  geom_polygon(data=lin.hull[lin.hull=="NA1",],
               aes(x=v.PCoA1,y=v.PCoA2),
               colour="black",alpha=0,linetype="dashed")+
  geom_segment(data=lin.data[lin.data$group == "NA1",],
               aes(x=v.PCoA1,xend=PCoA1,y=v.PCoA2,yend=PCoA2),
               alpha=0.30)+ 
  geom_point(data=L.cent[2,1:3],
             aes(x=PCoA1,y=PCoA2),
             size=4,colour="red",shape=17)+ 
  geom_point(data=lin.data[lin.data$group == "NA1",],
             aes(x=v.PCoA1,y=v.PCoA2),size=2,shape=17)+
  labs(title="B.",x="",y="")+
  coord_cartesian(xlim = c(-0.5,0.5), ylim = c(-0.5,0.5))+
  theme_bw()+ 
  theme(legend.position="none")
panel.L <- ggplot()+ 
  geom_polygon(data=lin.hull,
               aes(x=v.PCoA1,y=v.PCoA2),
               colour="black",alpha=0,linetype="dashed")+
  geom_segment(data=lin.data,
               aes(x=v.PCoA1,xend=PCoA1,y=v.PCoA2,yend=PCoA2),
               alpha=0.30)+ 
  geom_point(data=L.cent[,1:3],
             aes(x=PCoA1,y=PCoA2,shape=grps),
             size=4,colour="red")+ 
  geom_point(data=lin.data,
             aes(x=v.PCoA1,y=v.PCoA2,shape=group),size=2)+
  labs(title="C.",x="",y="")+
  coord_cartesian(xlim = c(-0.5,0.5), ylim = c(-0.5,0.5))+
  theme_bw()+ 
  theme(legend.position="none")

p.perL <- ggarrange(panel.eu,panel.na,panel.L,ncol = 3,nrow = 1)
p.perL

ggsave(filename = "Perm_Lin.png",
       p.perL,
       width = 6.5,
       height = 2,
       dpi = 300,
       units = "in",
       device = 'png')
```
