---
title: "Curry Co. SOD Analysis"
author: "Hazel Daniels"
date: "2/5/2021"
output:
  word_document: default
  html_document: default
---

# R setup and libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
library(readxl)
library(ggplot2)
library(vegan)
library(ape)
library(tidyverse)
library(viridis)
library(jtools)
library(car)
library(MASS)
library(gvlma)
library(nlme)
library(broom)
library(rstatix)
library(ggpubr)
library(emmeans)
```

# Reading and manipulating data
```{r, inlcude=FALSE}
SODSites <- read_xlsx("BasicData.xlsx", 1)
head(SODSites)
SODSite.ord <- SODSites[-c(31),,drop=F] #removing HD-4331 as it skews results strongly

# Subsetting data
AllSites <- SODSites[-c(76, 78, 79, 82, 83, 84, 88, 97, 98, 99),,drop=F] #removing HD-4376, 4378, 4379, 4382, 4383, 4384, 4388, 4397, 4398, 4399 as these sites were measured before and after treatment (removing treated data). Removing these duplicate sites to a data subset for separate analysis.
AllSites <- subset(AllSites, AllSites$Lineage != "U") #removing uninfested sites
AllSites$TrtBin <- as.factor(AllSites$TrtBin)

NA1 <- subset(AllSites, AllSites$Lineage == "NA1")

AllSeeds <- read_xlsx("BasicData.xlsx", 2)
AllSeeds <- column_to_rownames(AllSeeds, var = "New Site")
head(AllSeeds)
AllSeeds <- AllSeeds[-11] #removing CE as it skews results strongly
AllSeeds <- AllSeeds[-c(31),,drop=F] #removing HD-4331 as it skews results strongly

AllFire <- read_xlsx("BasicData.xlsx", 3)
AllFire$RX <- as.factor(AllFire$RX)
head(AllFire)

EU1wide <- read_xlsx("BasicData.xlsx", 4)
head(EU1wide)

EU1long <- read_xlsx("BasicData.xlsx", 5)
head(EU1long)
```

# Visualizing raw data
```{r}
ggplot(AllSites, aes(x=x)) +
  geom_histogram( aes(x = VegProp, y = stat(count))) +
  scale_x_continuous(name = "Was P.ramorum found in residual Veg sample?", breaks = seq(0, 1)) +
  facet_grid(Lineage ~ TrtBin) +
  labs(title = "Veg") +
  theme_bw()

ggplot(AllSites, aes(x=x)) +
  geom_histogram( aes(x = SoilProp, y = stat(count))) +
  scale_x_continuous(name = "Was P.ramorum found in residual Soil sample?", breaks = seq(0, 1)) +
  facet_grid(Lineage ~ TrtBin) +
  labs(title = "Soil") +
  theme_bw()

Trt <- c(rep("Untreated", 4), rep("Treated", 4))
Lin <- c("EU1","EU1","NA1","NA1","EU1","EU1","NA1","NA1")
Tis <- c("Vegetation","Soil","Vegetation","Soil","Vegetation","Soil","Vegetation","Soil")
Count <- c(54, 9, 51, 9, 15, 7, 28, 9)
df <- data.frame(Trt,Lin,Tis,Count)
df

ggplot(df,aes(x=Tis,y=Count,fill=Lin))+
  geom_bar(stat="identity",position="dodge")+
  scale_fill_discrete(name="Lineage",labels=c("EU1", "NA1"))+
  scale_fill_viridis_d() +
  theme_classic()+
  theme(legend.position="none", text = element_text(size=15))+
  xlab("Source material tested positive for P. ramorum")+
  ylab("Count")+
  ylim(0, 60)+
  geom_text(aes(label = Count), vjust = -0.5)+
  facet_grid(Lin~Trt)

#This visualization includes all 100 sites sampled, including uninfested sites, and EU1 sites that were sampled before and after treatment.
```


# Vegetation
## EU1 Repeated Measures
```{r}
V3 <- lme(VegProp ~ TrtBin, random = ~1|Plot, data = EU1long)

anova(V3)
summary(V3)

DTV <-array(c(1,8,0,1),
           dim=c(2, 2),
           dimnames=list(
             Time2=c("ThenPos", "ThenNeg"), 
             Time1=c("FirstPos", "FirstNeg"))
           )
DTV

mcnemar.test(DT, y=NULL, correct = FALSE)
chisq.test(EU1long$VegProp, EU1long$TrtBin)
fisher.test(EU1long$VegProp, EU1long$TrtBin)

#correlations
VC3 <- cor.test(EU1wide$VegPre, EU1wide$VegPost, method=c("pearson", "kendall", "spearman"))
VC3
```

## NA1 Veg Wildfire
```{r}
V2 <- lm(VegProp ~ RX, data = AllFire)

anova(V2)
summary(V2)
summ(V2)

# Contrast statement
levels(AllFire$RX)
# 0 - no WF, no RX
# 1 - no WF, yes RX
# 2 - yes WF, no RX
# 3 - yes WF, yes RX

V2.lsm <- lsmeans(V2, ~RX, adjst="tukey")
V2.lsm
plot(V2.lsm)

ConV2 <- contrast(V2.lsm, list (
  Ze.v.On = c(-1, 1, 0, 0), #difference is RX
  Tw.v.Th = c(0, 0, -1, 1), #difference is RX
  On.v.Tw = c(0, -1, 1, 0), #difference is both
  Th.v.Ze = c(1, 0, 0, -1), #difference is both
  Ze.v.Tw = c(-1, 0, 1, 0), #difference is WF; NS b/c small n & wide CI?
  On.v.Th = c(0, -1, 0, 1), #difference is WF
  OnTh.v.Tw = c(0, .5, -1, .5), #both RX to WF w/o RX
  OnTh.v.Ze = c(-1, .5, 0, .5), #both RX to noWF or RX
  TwTh.v.Ze = c(-1, 0, .5, .5), #both WF to noWF or RX
  TwTh.v.On = c(0, -1, .5, .5)  #both WF to RX w/o WF
  ))
ConV2

#correlations of random effects not included in the model
VC2 <- cor.test(AllFire$VegProp, AllFire$TotalTO, method=c("pearson", "kendall", "spearman"))
VC22 <- cor.test(AllFire$VegProp, AllFire$SoilProp, method=c("pearson", "kendall", "spearman"))

VC2
VC22 #correlation between soil and veg!
```

## NA1 Veg Wildfire Assumptions
```{r}
plot(V2)
gvlma(V2)

#constant variance
spreadLevelPlot(V2)

#normality of residuals
Resq2 <- qqPlot(V2, main="QQ Plot")

Res2 <- studres(V2)
hist(Res2, freq = FALSE,
     main="Distribution of Studentized Residuals")
xfit2 <- seq(min(Res2), max(Res2), length = 40)
yfit2 <- dnorm(xfit2)
lines(xfit2, yfit2)

#outliers
AllFire$Wleverage <- hatvalues(V2)
ggplot(AllFire, aes(`New Site`, Wleverage)) + geom_point() + ylim(0,1) + xlab("test")
```

## All Veg
```{r}
V1 <- lm(VegProp ~ Lineage * TrtBin, data = AllSites)

anova(V1)
summary(V1)
summ(V1)

# Contrast statement
V1.lsm <- lsmeans(V1, "TrtBin")
V1.lsm
plot(V1.lsm)

ConV1 <- contrast(V1.lsm, list (TvU = c(-1, 1)))
ConV1

#correlations of random effects not included in the model
VC1 <- cor.test(AllSites$VegProp, AllSites$YrSinceTrt, method=c("pearson", "kendall", "spearman"))
VC11 <- cor.test(AllSites$VegProp, AllSites$TotalTO, method=c("pearson", "kendall", "spearman"))
VC111 <- cor.test(AllSites$VegProp, AllSites$Num_Seedlings, method=c("pearson", "kendall", "spearman"))
VC1111 <- cor.test(AllSites$VegProp, AllSites$SoilProp, method=c("pearson", "kendall", "spearman"))
VC1
VC11
VC111
VC1111 #correlation between soil and veg!
```

## All Veg Assumptions
```{r}
plot(V1)
gvlma(V1)

#linearity
ggplot(AllSites, aes(TrtBin,VegProp)) + 
  stat_smooth(method="loess") + 
  stat_smooth(method="lm", color="red", fill="red", alpha=.25) + 
  geom_point()

#constant variance
spreadLevelPlot(V1)

#normality of residuals
Resq1 <- qqPlot(V1, main="QQ Plot")

Res1 <- studres(V1)
hist(Res1, freq = FALSE, main="Distribution of Studentized Residuals")
xfit1 <- seq(min(Res1), max(Res1), length = 40)
yfit1 <- dnorm(xfit1)
lines(xfit1, yfit1)

#multicollinearity
vif(V1)
sqrt(vif(V1)) > 2

#outliers
AllSites$Vleverage <- hatvalues(V1)
ggplot(AllSites, aes(Plot, Vleverage)) + geom_point() + ylim(0,1) + xlab("test")
```


# Soil
## EU1 Repeated Measures
```{r}
S3 <- lme(SoilBin ~ TrtBin, random = ~1|Plot, data = EU1long)

anova(S3)
summary(S3)

DTS <-array(c(1,4,1,4),
           dim=c(2, 2),
           dimnames=list(
             Time2=c("ThenPos", "ThenNeg"), 
             Time1=c("FirstPos", "FirstNeg"))
           )
DTS

mcnemar.test(DTS, y=NULL, correct = FALSE)
chisq.test(EU1long$SoilBin, EU1long$TrtBin)
fisher.test(EU1long$SoilBin, EU1long$TrtBin)

#correlations
SC3 <- cor.test(EU1wide$SoilPre, EU1wide$SoilPost, method=c("pearson", "kendall", "spearman"))
SC3
```

## NA1 Soil Wildfire
```{r}
S2 <- glm(SoilBin ~ RX, data = AllFire)

anova(S2)
summary(S2)
summ(S2)

# Contrast statement
levels(AllFire$RX)
# 0 - no WF, no RX
# 1 - no WF, yes RX
# 2 - yes WF, no RX
# 3 - yes WF, yes RX

S2.lsm <- lsmeans(S2, ~RX, adjst="tukey")
S2.lsm
plot(S2.lsm)

#correlations of random effects not included in the model
SC2 <- cor.test(AllFire$VegProp, AllFire$TotalTO, method=c("pearson", "kendall", "spearman"))
SC22 <- cor.test(AllFire$VegProp, AllFire$SoilProp, method=c("pearson", "kendall", "spearman"))

SC2
SC22 #correlation between soil and veg!
```

## NA1 Soil Wildfire Assumptions
```{r}
plot(S2)

#constant variance
spreadLevelPlot(S2)

#normality of residuals
Res4 <- studres(S2)
hist(Res4, freq = FALSE,
     main="Distribution of Studentized Residuals")
xfit4 <- seq(min(Res4), max(Res4), length = 40)
yfit4 <- dnorm(xfit4)
lines(xfit4, yfit4)

#outliers
AllFire$WSleverage <- hatvalues(S2)
ggplot(AllFire, aes(`New Site`, WSleverage)) + geom_point() + ylim(0,1) + xlab("test")
```

## All Soil
```{r}
S1 <- glm(SoilBin ~  Lineage + TrtBin, family = "binomial", data = AllSites)

anova(S1)
summary(S1)
summ(S1)

# Contrast statement
S1.lsm <- lsmeans(S1, "TrtBin")
S1.lsm
plot(S1.lsm)

#correlations of random effects not included in the model
SC1 <- cor.test(AllSites$SoilBin, AllSites$YrSinceTrt, method=c("pearson", "kendall", "spearman"))
SC11 <- cor.test(AllSites$SoilBin, AllSites$TotalTO, method=c("pearson", "kendall", "spearman"))
SC111 <- cor.test(AllSites$SoilBin, AllSites$Num_Seedlings, method=c("pearson", "kendall", "spearman"))
SC1111 <- cor.test(AllSites$SoilBin, AllSites$VegProp, method=c("pearson", "kendall", "spearman"))
SC1
SC11
SC111
SC1111 #correlation between soil and veg!
```

## All Soil Assumptions
```{r}
plot(S1)

#linearity
ggplot(AllSites, aes(SoilBin,TrtBin)) + 
  stat_smooth(method="loess") + 
  stat_smooth(method="lm", color="red", fill="red", alpha=.25) + 
  geom_point()

#constant varience
spreadLevelPlot(S1)

#normality of residuals
Res3 <- studres(S1)
hist(Res3, freq = FALSE,
     main="Distribution of Studentized Residuals")
xfit3 <- seq(min(Res3), max(Res3), length = 40)
yfit3 <- dnorm(xfit3)
lines(xfit3, yfit3)
#assumption violated?

#multicollinearity
vif(S1)
sqrt(vif(S1)) > 2

#outliers
AllSites$Sleverage <- hatvalues(S1)
ggplot(AllSites, aes(Plot, Sleverage)) + geom_point() + ylim(0,1) + xlab("test")
```


# Ordination of seedling data
## NMDS
```{r}
PCASeeds <- AllSeeds[5:35]
DistSeed <- vegdist(PCASeeds, method = "bray")

NMDS.scree <- function(x) { #where x is the name of the data frame variable
  plot(rep(1, 10), replicate(10, metaMDS(x, autotransform = F, k = 1)$stress), xlim = c(1, 10),ylim = c(0, 0.30), xlab = "# of Dimensions", ylab = "Stress", main = "NMDS stress plot")
  for (i in 1:10) {
    points(rep(i + 1,10),replicate(10, metaMDS(x, autotransform = F, k = i + 1)$stress))
  }
}

#NMDS.scree(DistSeed) chunk excluded to save space
```

```{r, include=FALSE}
NMDS.scree(DistSeed)
```

```{r}
NMDS1 <- metaMDS(DistSeed, k = 2, trymax = 100, trace = F)
NMDS1
Stress <- stressplot(NMDS1)
Stressplot <- plot(NMDS1, type = "t")

NMDS2 <- metaMDS(PCASeeds, k = 2, trymax = 100, trace = F)
NMDS2

NMDS3 <- metaMDS(PCASeeds, k = 2, trymax = 100, trace = F, autotransform = FALSE, distance="bray")
NMDS3plot <- plot(NMDS3)
NMDS3plotsp <- plot(NMDS3, display = "species")
```

## Interpretation (including effects variables)
```{r}
#Interpretation
EFvars <- c("New Site", "Lineage", "Latitude", "Longitude", "VegBin", "SoilBin", "YrSinceTrt")
EFSites <- SODSite.ord[EFvars]
head(EFSites)

ef <- envfit(NMDS3, EFSites, permutations = 999, na.rm = TRUE)

efplot <- plot(NMDS3, type = "t", display = "species")
efplot2 <- plot(ef, p.max = 0.05)
efplot2
```

## Interpretation by treatment
```{r}
group = c(rep("Untreated", 56), rep("Treated", 43))
colors = c(rep("red", 56), rep("blue", 43))

ordiplot(NMDS3, type = "n")
for(i in unique(group)) {
  ordihull(NMDS3$point[grep(i, group),], draw="polygon",
  groups = group[group == i],col = colors[grep(i,group)],label=F) }

orditorp(NMDS3, display = "species", col = "red", air = 0.01)
#orditorp(NMDS3, display = "sites", col = c(rep("red",12),
#  rep("blue", 12)), air = 0.01, cex = 1.25)
```

```{r}
#MRPP or other analysis coming soon!
```