---
title: "Radiata Inoculation"
author: "Hazel Daniels"
output:
  html_document:
    df_print: paged
updated: April 16, 2020
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Artificial inoculation
```{r}
# reading in file
library(readxl)
Inoc.data <- read_xlsx("C:/Users/dixie/Google Drive/_Ph.D. Forestry/Research/2018.11_RadiataInoculation/2019.01.14_LogInoculationDataSheet.xlsx", sheet=1)

#convert characters to numeric
Inoc.data$Inoc_1_m1_cm <- as.numeric(as.character(Inoc.data$Inoc_1_m1_cm))
Inoc.data$Inoc_1_m2 <- as.numeric(as.character(Inoc.data$Inoc_1_m2))
Inoc.data$Inoc_1_m3 <- as.numeric(as.character(Inoc.data$Inoc_1_m3))
Inoc.data$Inoc_1_m4 <- as.numeric(as.character(Inoc.data$Inoc_1_m4))
Inoc.data$Inoc_1_avg <- as.numeric(as.character(Inoc.data$Inoc_1_avg))
Inoc.data$Inoc_2_m1_cm <- as.numeric(as.character(Inoc.data$Inoc_2_m1_cm))
Inoc.data$Inoc_2_m2 <- as.numeric(as.character(Inoc.data$Inoc_2_m2))
Inoc.data$Inoc_2_m3 <- as.numeric(as.character(Inoc.data$Inoc_2_m3))
Inoc.data$Inoc_2_m4 <- as.numeric(as.character(Inoc.data$Inoc_2_m4))
Inoc.data$Inoc_2_avg <- as.numeric(as.character(Inoc.data$Inoc_2_avg))

Inoc.data$DIB_avg <- rowMeans(Inoc.data[c('DIB_1_cm', 'DIB_2')], na.rm=TRUE)

# data manipulation
Nden <- subset(Inoc.data, Species == "Nden")
Pmen <- subset(Inoc.data, Species == "Pmen")
Prad <- subset(Inoc.data, Species == "Prad")
EU1 <- subset(Inoc.data, Inoc_1 == "EU1")
NA1 <- subset(Inoc.data, Inoc_1 == "NA1")
```

## Is lesion length of artificial inoculation different between Species?
```{r}
# summary of inoculations
library(dplyr)
group_by(Inoc.data, Species) %>%
  summarise(
    count = n(),
    mean = mean(Inoc_1_avg, na.rm = TRUE),
    sd = sd(Inoc_1_avg, na.rm = TRUE),
    median = median(Inoc_1_avg, na.rm = TRUE)
  )

# visualize
library("ggpubr")
ggboxplot(Inoc.data, x = "Species", y = "Inoc_1_avg", 
          color = "Species", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          order = c("Nden", "Prad", "Pmen"),
          ylab = "Average lesion size (cm)", xlab = "Species")

ggline(Inoc.data, x = "Species", y = "Inoc_1_avg", 
       add = c("mean_se", "jitter"), 
       order = c("Nden", "Prad", "Pmen"),
       ylab = "Average lesion size (cm)", xlab = "Species")

# ANOVA
Inoc.anova <- aov(Inoc_1_avg ~ Species, data = Inoc.data)
summary(Inoc.anova)

# Tukey pairwise comparisons
TukeyHSD(Inoc.anova)

# testing assumption of equal variance
plot(Inoc.anova, 1)

library(car)
leveneTest(Inoc_1_avg ~ Species, data = Inoc.data)
#assumption violated!

# ANOVA w/o assumption of equal variance
oneway.test(Inoc_1_avg ~ Species, data = Inoc.data)

# pairwise comparison w/o pooled variance
pairwise.t.test(Inoc.data$Inoc_1_avg, Inoc.data$Species, p.adjust.method = "BH", pool.sd = FALSE)

# normality test
plot(Inoc.anova, 2)
#assumption valid
```

## Is lesion length of artificial inoculation different between lineage?
```{r}
group_by(Inoc.data, Inoc_1) %>%
  summarise(
    count = n(),
    mean = mean(Inoc_1_avg, na.rm = TRUE),
    sd = sd(Inoc_1_avg, na.rm = TRUE),
    median = median(Inoc_1_avg, na.rm = TRUE)
  )

# visualize
ggboxplot(Inoc.data, x = "Inoc_1", y = "Inoc_1_avg", 
          color = "Inoc_1", palette = c("#00AFBB", "#FC4E07"),
          order = c("EU1", "NA1"),
          ylab = "Average lesion size (cm)", xlab = "Lineage")

## by lineage by species
ggboxplot(Inoc.data, x = "Inoc_1", y = "Inoc_1_avg", 
          color = "Species", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          order = c("EU1", "NA1"),
          ylab = "Average lesion size (cm)", xlab = "Lineage")

ggline(Inoc.data, x = "Inoc_1", y = "Inoc_1_avg", 
       add = c("mean_se", "jitter"), 
       order = c("EU1", "NA1"),
       ylab = "Average lesion size (cm)", xlab = "Lineage")

# ANOVA
Lin.anova <- aov(Inoc_1_avg ~ Inoc_1, data = Inoc.data)
summary(Lin.anova)

# Tukey pairwise comparisons
TukeyHSD(Lin.anova)

# testing assumption of equal variance
plot(Lin.anova, 1)

library(car)
leveneTest(Inoc_1_avg ~ Inoc_1, data = Inoc.data)
#assumption violated!

# ANOVA w/o assumption of equal variance
oneway.test(Inoc_1_avg ~ Inoc_1, data = Inoc.data)

# pairwise comparison w/o pooled variance
pairwise.t.test(Inoc.data$Inoc_1_avg, Inoc.data$Inoc_1, p.adjust.method = "BH", pool.sd = FALSE)

# normality test
plot(Lin.anova, 2)
#assumption valid
```

## Is necrosis of control inoculation different between species?
```{r}
#summary of controls
group_by(Inoc.data, Species) %>%
  summarise(
    count = n(),
    mean = mean(Inoc_2_avg, na.rm = TRUE),
    sd = sd(Inoc_2_avg, na.rm = TRUE),
    median = median(Inoc_2_avg, na.rm = TRUE)
  )

# visualize
ggboxplot(Inoc.data, x = "Species", y = "Inoc_2_avg", 
          color = "Species", palette = c("#00AFBB", "#E7B800", "#FC4E07"),
          order = c("Nden", "Prad", "Pmen"),
          ylab = "Average reaction size (cm)", xlab = "Species")

ggline(Inoc.data, x = "Species", y = "Inoc_2_avg", 
       add = c("mean_se", "jitter"), 
       order = c("Nden", "Prad", "Pmen"),
       ylab = "Average reaction size (cm)", xlab = "Species")

# ANOVA
Ctrl.anova <- aov(Inoc_2_avg ~ Species, data = Inoc.data)
summary(Ctrl.anova)

# Tukey pairwise comparisons
TukeyHSD(Ctrl.anova)

# testing assumption of equal variance
plot(Ctrl.anova, 1)

library(car)
leveneTest(Inoc_2_avg ~ Species, data = Inoc.data)
#assumption violated!

# ANOVA w/o assumption of equal variance
oneway.test(Inoc_2_avg ~ Species, data = Inoc.data)

# pairwise comparison w/o pooled variance
pairwise.t.test(Inoc.data$Inoc_2_avg, Inoc.data$Species, p.adjust.method = "BH", pool.sd = FALSE)

# normality test
plot(Ctrl.anova, 2)
#assumption violated?

# Kruskal-Wallis
kruskal.test(Inoc_2_avg ~ Species, data = Inoc.data)
```

## Is lesion length of control inoculation different than arificial inoculation  within species?
```{r}
#radiata
t.test(Prad$Inoc_2_avg, Prad$Inoc_1_avg, var.equal = FALSE)

#tanoak
t.test(Nden$Inoc_2_avg, Nden$Inoc_1_avg, var.equal = FALSE)

#doug-fir
t.test(Pmen$Inoc_2_avg, Pmen$Inoc_1_avg, var.equal = FALSE)
```

## Is reisolation (e.g. Culture = Y) predicted by bark depth, diameter (DIB), P. ramorum lineage, and/or species?
```{r}
Cult_null <- glm(CultBin~1, family = binomial, data=Inoc.data)
Cult_full <- glm(CultBin~Species*DIB_avg*Bark*Inoc_1, family=binomial, data=Inoc.data)
step(Cult_null, scope = list(upper=Cult_full), direction="both", test="Chisq", data=Inoc.data)

Inoc.Cult <- glm(formula = CultBin~1+Species+DIB_avg+Species:DIB_avg, family = binomial, data = Inoc.data)

summary(Inoc.Cult)
```

## Are reisolation rates different between species?
```{r}
# create contingency table
RIR.table <- matrix (c(10, 1, 2, 8, 15, 2), nrow = 3, ncol = 2, byrow = TRUE)
rownames(RIR.table) <- c("Nden", "Pmen", "Prad")
colnames(RIR.table) <- c("Pos", "Neg")
RIR.table

# chi-square analysis
RIR.chi <- chisq.test(RIR.table)
RIR.chi
RIR.chi.sim <- chisq.test(RIR.table, simulate.p.value = TRUE)
RIR.chi.sim
```

## Are reisolation rates different between lineage?
```{r}
# create contingency table
RIR.tab.lin <- matrix (c(14, 13, 5, 6), nrow = 2, ncol = 2)
rownames(RIR.tab.lin) <- c("EU1", "NA1")
colnames(RIR.tab.lin) <- c("Pos", "Neg")
RIR.tab.lin

# chi-square analysis
RIR.chi.lin <- chisq.test(RIR.tab.lin)
RIR.chi.lin
RIR.chi.lin.sim <- chisq.test(RIR.tab.lin, simulate.p.value = TRUE)
RIR.chi.lin.sim
```

# Field Inoculation
```{r}
# reading in file
Field.data <- read_xlsx("C:/Users/dixie/Google Drive/_Ph.D. Forestry/Research/2018.11_RadiataInoculation/2019.01.14_LogInoculationDataSheet.xlsx", sheet=2)
Field.data
Field.data$Site <- as.character(as.numeric(Field.data$Site))

# averaging DIBs
Field.data$DIB_avg <- rowMeans(Field.data[c('DIB_1_cm', 'DIB_2')], na.rm=TRUE)

summary(Field.data)

# data manipulation
Field_Nden <- subset(Field.data, Species == "Nden")
Field_Pmen <- subset(Field.data, Species == "Pmen")
Field_Prad <- subset(Field.data, Species == "Prad")
```

## Is isolation (e.g. Culture = Y) predicted by site, bark depth, diameter (DIB), and/or species?
```{r}
Field_null <- glm(CultBin~1, family = binomial, data=Field.data)
Field_full <- glm(CultBin~Site*Species*DIB_avg*Bark, family=binomial, data=Field.data)
step(Field_null, scope = list(upper=Field_full), direction="both", test="Chisq", data=Field.data)

Field.Cult <- glm(formula = CultBin ~ DIB_avg + Bark, family = binomial, data = Field.data)

summary(Field.Cult)
```
